biclaR: Estimating the socio-environmental impacts of car substitution by bicycle and public transit using open tools

### 2023-10-03 paper/PaperTRA/PaperTRA.Rmd
correct summary 2 and 3

summary21 = summary21 %>%
  # mutate(LTS = recode(LTS, "seguro" = "safe", "direto" = "direct")) |> 
  # mutate(Estrategia = paste0(`Meta ENMAC`*100,"%")) |> 
  # select(Estrategia, LTS, Total, Autocarro, Barco, Comboio, `Metro superfície`) |> 
  arrange(desc(Target))

@@

```{r summary22}

summary22 = summary2 %>%
  mutate(LTS = recode(LTS, "seguro" = "safe", "direto" = "direct")) |> 
  mutate(Estrategia = paste0(`Meta ENMAC`*100,"%")) |> 
  mutate(Value = `Valorização monetária (€)`/1000) |> 
  select(Estrategia, LTS, CO2eq, CO, PM10, NOx, VOC, Value) |> 
  arrange(desc(Estrategia))

@@

summaryall = summary1 |> left_join(summary22) |> 
  mutate(Code = c(3304, 3404, 3310, 3410)) |> 
  left_join(summary0 |> select(Code, Economic), by = "Code") |> 
  select(c(1, 2, 6, 7, 8, 13, 15)) |> 
  mutate(CO2eq = co2 + CO2eq,
         value1 = Economic/1000 + Value, # this is CORRECT
         value10 = monet + Value*10*rate) |> # this is CORRECT
  select(Estrategia, LTS, CO2eq, value1, value10)

>>>>

summary21 = summary2 %>%
  # mutate(LTS = recode(LTS, "seguro" = "safe", "direto" = "direct")) |> 
  # mutate(Estrategia = paste0(`Meta ENMAC`*100,"%")) |> 
  # select(Estrategia, LTS, Total, Autocarro, Barco, Comboio, `Metro superfície`) |> 
  arrange(desc(Target))

@@

```{r summary22}
summary22 = readRDS("load/TPtable_aml.Rds")
summary22 = summary22 %>%
  mutate(LTS = recode(Percursos, "seguro" = "safe", "direto" = "direct")) |> 
  mutate(Value = money_emissions_eur/1000) |> 
  select(Meta_ENMAC, LTS, CO2eq_TP, CO_PT, PM10_PT, NOx_PT, VOC_PT, Value) |> 
  arrange(Meta_ENMAC)

@@

summaryall = summary1 |> 
  mutate(Meta_ENMAC = Estrategia) |>
  left_join(summary22) |> 
  mutate(Code = c(3304, 3404, 3310, 3410)) |> 
  left_join(summary0 |> select(Code, Economic), by = "Code") |> 
  select(Estrategia, LTS, co2, CO2eq_TP, monet, Value, Economic) |> 
  mutate(CO2eq = co2 + CO2eq_TP,
         value1 = Economic/1000 + Value, # this is CORRECT
         value10 = monet + Value*10*rate) |> # this is CORRECT
  select(Estrategia, LTS, CO2eq, value1, value10)

### 2023-10-03 paper/PaperTRA/PaperTRA.tex
correct summary 2 and 3

The \texttt{r5r} model used the OpenStreetMap road network and the GTFS
metropolitan data agregated and validated. This information is crucial
for an accurate PT trip and route estimation. A digital elevation model
from the European Space Agency's COPERNICUS mission, with a 25m spatial
resolution, was used to include street gradient information, as a weight
in cycling routing. The cycling potential trips for the two national
strategic targets (4\% and 10\%) was estimated from the values for
cycling and car trips (both as a driver and as a passenger) from the
2017 base scenario.

@@

For the cycling legs of the journey (first and last legs),
socio-environmental impacts were estimated, using the HEAT for Cycling
tool v5.0 {[}12{]} and the
\href{https://github.com/HEAT-WHO/HEAT_heatr_api}{\texttt{HEATaaS} R
package}\footnote{\texttt{HEATaaS} is under development. For more
  information contact
  \href{https://heatwalkingcycling.org}{heatwalkingcycling.org}.} from
the World Health Organization. The HEAT tool provided estimates on the
shifting from car to cycling for a short term time horizon (i.e., one
year) and the long term (i.e., ten years). We considered two dimensions:
\emph{social} --- including the physical activity of cyclists, air
pollution exposure, and road casualties; \emph{environmental} ---
including CO\textsubscript{2}eq emissions and other pollutants.

@@

For both \emph{direct} and \emph{safe} route profiles, 20\% of the daily
car trips have the potential to be replaced by a combination of PT and
cycling (up to 5 km on bike). \emph{Table \ref{tab:summary21} shows
something that doesn't make sense.}

@@

\caption{\label{tab:summary21}\label{summary21}Summary of the potential of replacing car trips with cycling in combination with PT, disagregated by PT mode.}
\centering
\begin{tabular}[t]{ll>{\raggedleft\arraybackslash}p{4em}>{\raggedleft\arraybackslash}p{4em}>{\raggedleft\arraybackslash}p{4em}>{\raggedleft\arraybackslash}p{4em}>{\raggedleft\arraybackslash}p{4em}}
\toprule
Target & Routing & Total & Bus & Ferry & Train & Tram\\
\midrule
4\% & safe & 192 214 & 4 163 & 5 167 & 169 984 & 12 900\\
4\% & direct & 189 846 & 4 879 & 5 795 & 171 534 & 7 639\\
10\% & safe & 224 152 & 5 042 & 5 594 & 197 857 & 15 659\\
10\% & direct & 219 511 & 5 806 & 6 263 & 198 373 & 9 069\\
\bottomrule
\end{tabular}
\end{table}

This scenario unveils the potential of cycling as a complementary mode
of PT, with the potential to uptake the number of PT trips within the
LMA area by as much as 12\% (in addition to the 825 thousand PT trips
reported in the mobility survey). These findings indicate that
transferring car trips to a combination of bicycle and PT could be
nearly as substantial, if not equally significant, as the shift towards
bicycle-only trips.

@@

The emissions of CO\textsubscript{2}eq that are avoided during both the
initial and final journey segments account for about 75\% of the
emissions avoided during the PT segment. This finding, while expected --
due the zero cycling emissions, should not be overlooked when promoting
the PT use. Improving the safe accessibility to PT interfaces to
cyclists and providing bicycle-friendly amenities such as parking
facilities can potentially lead to a higher reduction in
CO\textsubscript{2}eq emissions, compared to a scenario where
individuals shift from car travel to car + PT combination.

The social impacts from replacing car trips to bicycle in first-and-last
legs, represent 95\% of the total socio-environmental benefits estimated
for all journey legs. Our findings suggest that a bicycle and PT
combination could viably replace 20\% of current LMA trips, with an
additional 12\% of PT journeys prone to further substitution.

>>>>

The \texttt{r5r} model used the OpenStreetMap road network and the GTFS
metropolitan data aggregated and validated. This information is crucial
for an accurate PT trip and route estimation. A digital elevation model,
from the European Space Agency's COPERNICUS mission, was used to include
street gradient information, as a weight in cycling routing. The cycling
potential trips for the two national strategic targets (4\% and 10\%)
were estimated from the 2017 cycling and car trips (both as a driver and
as a passenger), the baseline scenario.

@@

For the cycling legs of the journey (first and last legs),
socio-environmental impacts were estimated, using the HEAT for Cycling
tool v5.0 {[}12{]} from the World Health Organization, and the
\href{https://github.com/HEAT-WHO/HEAT_heatr_api}{\texttt{HEATaaS} R
package}\footnote{\texttt{HEATaaS} is under development. For more
  information contact
  \href{https://heatwalkingcycling.org}{heatwalkingcycling.org}.}. The
HEAT tool provided estimates on the shifting from car to cycling for a
short term time horizon (i.e., one year) and the long term (i.e., ten
years). We considered two dimensions: \emph{social} --- including the
physical activity of cyclists, air pollution exposure, and road
casualties; \emph{environmental} --- including CO\textsubscript{2}eq
emissions and other pollutants.

@@

For both \emph{direct} and \emph{safe} route profiles, 20\% of the daily
car trips have the potential to be replaced by a combination of PT and
cycling (up to 5 km on bike). Table \ref{tab:summary21} shows the PT
modes with most potential of in combination with cycling, by mode.

@@

\caption{\label{tab:summary21}\label{summary21}Summary of the potential of replacing car trips with cycling in combination with PT, disagregated by PT mode.}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4.5em}>{\raggedright\arraybackslash}p{4.5em}>{\raggedleft\arraybackslash}p{4.5em}>{\raggedleft\arraybackslash}p{4.5em}>{\raggedleft\arraybackslash}p{4.5em}>{\raggedleft\arraybackslash}p{4.5em}>{\raggedleft\arraybackslash}p{4.5em}}
\toprule
Target & Routing & Potential & Bus & Ferry & Train & Tram\\
\midrule
4\% & safe & 20 385 & 573 & 285 & 17 716 & 1 811\\
4\% & direct & 18 944 & 593 & 313 & 17 093 & 946\\
10\% & safe & 52 323 & 1 452 & 712 & 45 588 & 4 571\\
10\% & direct & 48 609 & 1 520 & 781 & 43 932 & 2 375\\
\bottomrule
\end{tabular}
\end{table}

This scenario unveils the potential of cycling as a complementary mode
of PT, with the potential to uptake the number of PT trips within the
LMA area by as much as 6.3\% (in addition to the 825 thousand PT trips
reported in the mobility survey).

@@

The social impacts represent 98\% of the socio-environmental benefits
(in value) from replacing car trips to bicycle in first-and-last legs.
The emissions of CO\textsubscript{2}eq that are avoided during both the
initial and final journey segments account for about 75\% of the
emissions avoided during the PT segment. This finding, while expected --
due the zero cycling emissions, should not be overlooked when promoting
the PT use. Improving the safe accessibility to PT interfaces to
cyclists and providing bicycle-friendly amenities such as parking
facilities can potentially lead to a higher reduction in
CO\textsubscript{2}eq emissions, compared to a scenario where
individuals shift from car travel to car + PT combination.

Our findings suggest that a bicycle and PT combination could viably
replace 20\% of current LMA trips, with an additional 12\% of PT
journeys prone to further substitution.
